---
title: "HW2"
author: "Bulun Te"
date: "2024-01-26"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=TRUE}
library(dplyr)
library(survival)
library(kableExtra)
library(knitr)

# Reading the data

data = read.table("Breast_cancer_Table_1_2_Collet.txt")

# Changing the column names

data = data %>% rename(
  id = V1,
  x = V2,
  delta = V3,
  z = V4
)

data %>% head()

data %>% psych::describe()

```


```{r}

#data

#data$not_censored_x = as.numeric(data$x * data$delta>0)


# sort data with x column with aescending order

data_sorted = data %>% arrange(x)

data_sorted

data_sorted$y = sapply(data_sorted$x,function(u){sum(data_sorted$x >= u)})

data_sorted$lambda = data_sorted$delta / data_sorted$y

data_sorted$cumulative_hazard = cumsum(data_sorted$lambda)

cumulative_hazard = data_sorted %>% select(x,delta,y,lambda,cumulative_hazard) %>% filter(delta ==1) %>% select(cumulative_hazard)

ppl_at_risk = data_sorted %>% select(x,delta,y,lambda,cumulative_hazard) %>% filter(delta ==1) %>% select(x,y)

data2 = data.frame(time = ppl_at_risk$x, 
                    Y = ppl_at_risk$y,
                   cumulative_hazard= cumulative_hazard)

Surv = exp(-cumulative_hazard)

Surv

```






## Question 1.(a)

```{r}

obj = Surv(data$x, data$delta)

nelson_aalen_fit = survfit(obj ~ 1,type="fh")

plot(nelson_aalen_fit, cumhaz=TRUE, xlab = "Time", ylab = "Cumulative Hazard", main = "Nelson-Aalen Estimate")

result_frame = ((nelson_aalen_fit %>% summary())[c("time","surv")]) %>% as.data.frame() %>% mutate(time=time,Cumulative_Hazard = -log(surv))

result_frame

```


```{r}

obj = Surv(data$x, data$delta)

nelson_aalen_fit = survfit(obj ~ 1,type="fh",conf.type="log")

plot(nelson_aalen_fit, cumhaz=TRUE, 
     xlab = "Time", 
     ylab = "Cumulative Hazard", 
     main = "Nelson-Aalen Estimate",
     conf.int=FALSE)


```

## Question 1.(b)

Assuming the $\Lambda (t)$ follows normal distribution, then the confidence interval is estimated as

```{r}

obj = Surv(data$x, data$delta)

nelson_aalen_fit = survfit(obj ~ 1,type="fh",conf.type="log")

plot(nelson_aalen_fit, cumhaz=TRUE, xlab = "Time", ylab = "Cumulative Hazard", main = "Nelson-Aalen Estimate")

((nelson_aalen_fit %>% summary())[c("time","surv","lower","upper")])%>% 
  as.data.frame()%>% 
  mutate(time=time,
           Cumulative_Hazard = -log(surv),
           upper_lambda = -log(lower),
           lower_lambda = -log(upper)) %>%
  select(time,Cumulative_Hazard,lower_lambda,upper_lambda) %>% round(.,4)

```


## Question 1.(c)

Assuming the $log(\Lambda (t))$ follows normal distribution, then the confidence interval is estimated as

```{r}

obj = Surv(data$x, data$delta)

nelson_aalen_fit = survfit(obj ~ 1,type="fh",conf.type="log-log")

plot(nelson_aalen_fit, cumhaz=TRUE, xlab = "Time", ylab = "Cumulative Hazard", main = "Nelson-Aalen Estimate")

((nelson_aalen_fit %>% summary())[c("time","surv","lower","upper")])%>% 
  as.data.frame()%>% 
  mutate(time=time,
           Cumulative_Hazard = -log(surv),
           upper_lambda = -log(lower),
           lower_lambda = -log(upper)) %>%
  select(time,Cumulative_Hazard,lower_lambda,upper_lambda) %>% round(.,4)

```




## Question 1.(d)



```{r}

obj = Surv(data$x, data$delta)
nelson_aalen_survival = survfit(obj ~ 1,type="fh",conf.type="log-log")
nelson_aalen_survival %>% summary()

# plotting the survival function

plot(nelson_aalen_survival, xlab = "Time", ylab = "Survival Function", main = "Nelson-Aalen Estimate")

```

## Question 1.(e)

Find the median and quartiles of survival estimation

As the estimated minimum value of $S(t)$ is 0.358, the estimation of 0.75 quartile is unable to obtain


```{r}


obj = Surv(data$x, data$delta)

nelson_aalen_fit = survfit(obj ~ 1,type="fh",conf.type="log-log")

result = nelson_aalen_fit %>% summary()

print("quartiles of 0.25,0.5 of survival estimations are")

c(min(result$time[which(result$surv <= (1-0.25))]),
min(result$time[which(result$surv <= 0.5)]))

```
