---
title: "HW2"
author: "Bulun Te"
date: "2024-01-26"
output:
  pdf_document: 
    latex_engine: xelatex
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r,echo=TRUE,message=FALSE}
library(dplyr)
library(survival)
library(kableExtra)
library(knitr)
library(ggplot2)

# Reading the data

data = read.table("Breast_cancer_Table_1_2_Collet.txt")

# Changing the column names

data = data %>% rename(
  id = V1,
  x = V2,
  delta = V3,
  z = V4
)

data %>% head()

data %>% psych::describe()

```

# Answer without using survival package

## Question 1(a)

$\hat{\Lambda}_{NA}(t)=\sum_{t_j \le t} \frac{D_j}{Y_j}$

where $D_j$ is the number of events at time $t_j$ and 
$Y_j$ is the number of individuals at risk at time $t_j$.



```{r}


# sort data with x column with aescending order

data_sorted = data %>% arrange(x)

data_sorted

data_sorted$y = sapply(data_sorted$x,function(u){sum(data_sorted$x >= u)})

data_sorted$lambda = data_sorted$delta / data_sorted$y

data_sorted$cumulative_hazard = cumsum(data_sorted$lambda)

cumulative_hazard = data_sorted %>% 
  select(x,delta,y,lambda,cumulative_hazard) %>% 
  filter(delta ==1) %>% 
  select(cumulative_hazard)

ppl_at_risk = data_sorted %>% 
  select(x,delta,y,lambda,cumulative_hazard) %>% 
  filter(delta >=1) %>% select(x,y)

data2 = data.frame(time = ppl_at_risk$x, 
                    Y = ppl_at_risk$y,
                   Nelson_Alan_Cumaltive= cumulative_hazard)

data2
```


## Question 1(b)

Assuming the $\hat{\Lambda} (t)$ follows normal distribution, and the variance of $\hat{\Lambda}(t)$ is estimated as $\hat{V}(t)=\sum_{t_j \le t}\frac{D_j}{Y_j^2}$, then the confidence interval is estimated as $\hat{\Lambda}(t) \pm z_{0.975}\sqrt{\hat{V}(t)}$.



```{r}

# Computing 95% confidence interval for the cumulative hazard assuming lambda(t) follows normal distribution


V_estimate = ((data_sorted$delta)/(data_sorted$y^2)) %>% cumsum()

CI_upper = data_sorted$cumulative_hazard + qnorm(0.975)*sqrt(V_estimate)

CI_lower = data_sorted$cumulative_hazard - qnorm(0.975)*sqrt(V_estimate)

result_1_b = data_sorted %>% mutate(CI_upper = CI_upper, CI_lower = CI_lower)%>% 
  filter(delta>=1) %>% 
  select(x,cumulative_hazard,CI_upper,CI_lower) %>% 
  round(.,6)

result_1_b

ggplot(result_1_b, aes(x = x)) +
  geom_step(aes(y = cumulative_hazard), direction = "hv", col = "blue") +
  geom_ribbon(aes(ymin = CI_lower, ymax = CI_upper), alpha = 0.2) +
  labs(x = "Time", y = "Cumulative HAzard Nelson Alaan") +
  ggtitle("Nelson Alaan estimator with 95% Confidence Intervals") +
  theme_minimal()



```


## Question 1(c)

Assuming the $\hat{\Lambda} (t)$ follows normal distribution.  The variance of $\hat{\Lambda (t)}$ , using Delta method is estimated as $\hat{V}(\log(\hat{\Lambda} (t)) = \sum_{t_j \le t}\frac{D_j}{Y_j^2 \hat{\Lambda}(t)^2}$, then the confidence interval is estimated as $\exp(\log(\hat{\Lambda}(t) \pm z_{0.975}\sqrt{\hat{V}(\log(\hat{\Lambda}(t)))})$.

```{r}


# Computing 95% confidence interval for the cumulative hazard assuming 
# log(lambda(t)) follows normal distribution

# Computing using delta mathod

V_estimate = ((data_sorted$delta)/(data_sorted$y^2)) %>% cumsum()

V_estimate = V_estimate / (data_sorted$cumulative_hazard^2)

CI_upper = data_sorted$cumulative_hazard * exp(qnorm(0.975)*sqrt(V_estimate))

CI_lower = data_sorted$cumulative_hazard * exp(-qnorm(0.975)*sqrt(V_estimate))

result_1_c = data_sorted %>% 
  mutate(CI_upper = CI_upper, CI_lower = CI_lower)%>% 
  filter(delta>=1) %>% 
  select(x,cumulative_hazard,CI_upper,CI_lower) %>% 
  round(.,6)

result_1_c

ggplot(result_1_c, aes(x = x)) +
  geom_step(aes(y = cumulative_hazard), direction = "hv", col = "blue") +
  geom_ribbon(aes(ymin = CI_lower, ymax = CI_upper), alpha = 0.2) +
  labs(x = "Time", y = "Cumulative HAzard Nelson Alaan") +
  ggtitle("Nelson Alaan estimator with 95% Confidence Intervals") +
  theme_minimal()


```

## Question 1(d)

Assuming the $log(\Lambda (t))$ follows normal distribution.  The variance of $log(\Lambda (t))$ is estimated as above. As $S(t)=exp(-\Lambda(t))$, thus $P\{ \hat{\Lambda}(t) \in (a,b) \} = P\{ \hat{S}(t) \in (e^{-b},e^{-a}) \}=0.95$ And the CI are computed as follow:

```{r}

V_estimate_temp = ((data_sorted$delta)/(data_sorted$y^2)) %>% cumsum()

V_estimate = V_estimate_temp / (data_sorted$cumulative_hazard^2)

S_estimate = exp(-data_sorted$cumulative_hazard)

CI_lower = exp(-data_sorted$cumulative_hazard * exp(qnorm(0.975)*sqrt(V_estimate))) 

CI_upper = exp(-data_sorted$cumulative_hazard * exp(-qnorm(0.975)*sqrt(V_estimate))) 

result_1_d = data_sorted %>% 
  mutate(CI_upper = CI_upper, CI_lower = CI_lower,S_estimate=S_estimate)%>% 
  filter(delta>=1) %>% 
  select(x,CI_upper,S_estimate,CI_lower) %>%
  round(.,6)

result_1_d

# plotting the survival function and its confidence interval from result_1_d as step function

ggplot(result_1_d, aes(x = x)) +
  geom_step(aes(y = S_estimate), direction = "hv", col = "blue") +
  geom_ribbon(aes(ymin = CI_lower, ymax = CI_upper), alpha = 0.2) +
  labs(x = "Time", y = "Survival Probability") +
  ggtitle("Survival Function with 95% Confidence Intervals") +
  theme_minimal()


```

## Question 1(e)

Quartile of 0.75 is unable to compute due to the estimation of survival function ends in 0.358. However, the quartiles of 0.25 and 0.5 are computed as follow:

```{r}

print("quartiles of 0.25,0.5 of survival estimations are")

c(min(result_1_d$x[which(result_1_d$S_estimate <= (1-0.25))]),
min(result_1_d$x[which(result_1_d$S_estimate <= 0.5)]))

```


