---
title: "HW2_survival_package"
author: "Bulun Te"
date: "2024-02-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Answering with package survival

## Question 1.(a)

```{r}

obj = Surv(data$x, data$delta)

nelson_aalen_fit = survfit(obj ~ 1,type="fh")

plot(nelson_aalen_fit, cumhaz=TRUE, xlab = "Time", ylab = "Cumulative Hazard", main = "Nelson-Aalen Estimate")

result_frame = ((nelson_aalen_fit %>% summary())[c("time","surv")]) %>% as.data.frame() %>% mutate(time=time,Cumulative_Hazard = -log(surv))

result_frame

```


```{r}

obj = Surv(data$x, data$delta)

nelson_aalen_fit = survfit(obj ~ 1,type="fh",conf.type="log")

plot(nelson_aalen_fit, cumhaz=TRUE, 
     xlab = "Time", 
     ylab = "Cumulative Hazard", 
     main = "Nelson-Aalen Estimate",
     conf.int=FALSE)


```

## Question 1.(b)

Assuming the $\Lambda (t)$ follows normal distribution, then the confidence interval is estimated as

```{r}

obj = Surv(data$x, data$delta)

nelson_aalen_fit = survfit(obj ~ 1,type="fh",conf.type="log")

plot(nelson_aalen_fit, cumhaz=TRUE, xlab = "Time", ylab = "Cumulative Hazard", main = "Nelson-Aalen Estimate")

((nelson_aalen_fit %>% summary())[c("time","surv","lower","upper")])%>% 
  as.data.frame()%>% 
  mutate(time=time,
           Cumulative_Hazard = -log(surv),
           upper_lambda = -log(lower),
           lower_lambda = -log(upper)) %>%
  select(time,Cumulative_Hazard,lower_lambda,upper_lambda) %>% round(.,4)

```


## Question 1.(c)

Assuming the $log(\Lambda (t))$ follows normal distribution, then the confidence interval is estimated as

```{r}

obj = Surv(data$x, data$delta)

nelson_aalen_fit = survfit(obj ~ 1,type="fh",conf.type="log-log")

plot(nelson_aalen_fit, cumhaz=TRUE, xlab = "Time", ylab = "Cumulative Hazard", main = "Nelson-Aalen Estimate")

((nelson_aalen_fit %>% summary())[c("time","surv","lower","upper")])%>% 
  as.data.frame()%>% 
  mutate(time=time,
           Cumulative_Hazard = -log(surv),
           upper_lambda = -log(lower),
           lower_lambda = -log(upper)) %>%
  select(time,Cumulative_Hazard,lower_lambda,upper_lambda) %>% round(.,4)

```




## Question 1.(d)



```{r}

obj = Surv(data$x, data$delta)
nelson_aalen_survival = survfit(obj ~ 1,type="fh",conf.type="log-log")
result = nelson_aalen_survival %>% summary() 

result[c("time","upper","surv","lower")] %>% as.data.frame() %>% round(.,6)

# plotting the survival function

plot(nelson_aalen_survival, xlab = "Time", ylab = "Survival Function", main = "Nelson-Aalen Estimate")

```

## Question 1.(e)

Find the median and quartiles of survival estimation

As the estimated minimum value of $S(t)$ is 0.358, the estimation of 0.75 quartile is unable to obtain


```{r}


obj = Surv(data$x, data$delta)

nelson_aalen_fit = survfit(obj ~ 1,type="fh",conf.type="log-log")

result = nelson_aalen_fit %>% summary()

print("quartiles of 0.25,0.5 of survival estimations are")

c(min(result$time[which(result$surv <= (1-0.25))]),
min(result$time[which(result$surv <= 0.5)]))

```

